---
title: "Theming with bslib and thematic"
output: 
  html_document:
    code_folding: show
    theme:
      bg: "#202123"
      fg: "#B8BCC2"
      primary: "#EA80FC"
      secondary: "#00DAC6"
      base_font:
        google: Prompt
      heading_font:
        google: Proza Libre
---

```{r setup, include=FALSE}
if (requireNamespace("thematic")) 
  thematic::thematic_rmd(font = "auto")

knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```
## Loading libraries
```{r results='hide', message=FALSE, warning=FALSE}
library(fpc)
library(jpeg)
library(imager)
library(fields)
library(ggplot2)
library(raster)
library(dplyr)
library(cluster)
library(factoextra)
library(flexclust)
library(fpc)
library(ClusterR)
library(meanShiftR)
library(RColorBrewer)
library(Dict)
library(fdm2id)
library(showtext)
```
## Overview

This is an assignment created for Unsupervised Learning class. We were supposed to use clustering methods in order to analyze phenomena that are interesting for us.

**My project is answering the question:** Are we able to detect animals photographed by a trail camera using various clustering methods?

I will try to point out the animal seen on the photo by using couple of clustering methods, let's see if I will get even close to highlight only the animal, without any additional background.

The images are loaded in grayscale because this is the format most often registered by a trail camera. Because of the shade and darkness existing in the woods, sensor in the camera detects the low level of light and turns on the IR flash. The result is a grayscale picture.

Used methods: CLARA, DBSCAN, MEAN SHIFT, SPECTRAL CLUSTERING

Code and images are available on my [GitHub](https://github.com/dkoterwa/masters/tree/main/first%20semester/Unsupervised%20Learning/article_1)


## Functions {.tabset .tabset-pills}

I've created 4 functions to facilitate the work and create a clearer output at the end.

### Pick colours
```{r}

pick_colours = function (n) {
  
  #DESCRIPTION : this function is picking n contrasting colours from the palette
  #INPUT: n: number of colours that we want to pick
  #OUTPUT: list_of_colors: list of n colours
  
  #load palette
  qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
  
  #create colors vector
  col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
  
  #sample n colors from the vector
  list_of_colors = sample(col_vector, n, replace=FALSE)
  
  return (list_of_colors)
  
}
```

### Contrast image
```{r}
contrast_image = function (dataframe, scalar_white, scalar_black){
  
  #DESCRIPTION : this function is contrasting pixels of the grayscale image
  #INPUT: dataframe: 2D dataframe of our image
          #scalar_white: value which corresponds to how much we want to brighten light pixels
          #scalar_black: value which corresponds to how much we want to darken dark pixels
  #OUTPUT: dataframe: converted dataframe with contrasted pixels
  
  #if grayscale value is larger than a threshold - set a pixel value to scalar_white * pixel, else set it to scalar_black * pixel
  for (j in 1:ncol(dataframe)){
    for (i in 1:nrow(dataframe)){
      if(dataframe[i, j] > 0.3){
        dataframe[i, j] = dataframe[i,j] * scalar_white
      }
      else{
        dataframe[i, j] = dataframe[i, j] * scalar_black
      }
    }
  }
  return(dataframe)
}
```
### Pick colors for DBSCAN
```{r}
draw_colours_dbscan = function(dataframe, n_clusters){
  
  #DESCRIPTION : this function is assigning colours to specific clusters. It is made for DBSCAN algorithm,                  because this method can output cluster #0
  #INPUT: dataframe: 2D dataframe of our grayscale image with additional column 'cluster', which is the               assignment of each pixel to specific cluster
          #n_clusters: number of clusters which our algorithm has outputted
  #OUTPUT: dataframe: inputted dataframe with additional column 'color'
  
  if (n_clusters < 2){
    print ("Error: insufficient number of clusters : check your parameters")
    break}
    
    
  #sample n colours
  colours = pick_colours(n_clusters - 1)
  
  #compute number of clusters
  clusters = seq(from=1, to=n_clusters - 1, by=1)
  names(colours) = clusters
  
  #this is specific for dbscan: we want to assign color black to non-clustered pixels and contrasting colors for clusters
  for (i in 1:nrow(dataframe)){
    
    if(dataframe[i, ]['cluster'] != 0){
      dataframe[i, ]['color'] = colours[as.numeric(dataframe[i, ]['cluster'])]
    } else{
      dataframe[i , ]['color'] = '#000000'
    }
    
  }
  
  return(dataframe)
}

```
### Pick colors, other methods
```{r}
draw_colours = function(dataframe, n_clusters){
  
  #DESCRIPTION : this function is assigning colours to specific clusters. 
  #INPUT: dataframe: 2D dataframe of our grayscale image with additional column 'cluster', which is the               assignment of each pixel to specific cluster
          #n_clusters: number of clusters which our algorithm has outputted
  #OUTPUT: dataframe: inputted dataframe with additional column 'color'
  
  if (n_clusters < 2){
    print ("Error: insufficient number of clusters : check your parameters")
    break}
  
  
  #sample n colors
  colours = pick_colours(n_clusters)
  
  #compute number of clusters
  clusters = seq(from=1, to=n_clusters, by=1)
  names(colours) = clusters
  
  #this is specific for dbscan: we want to assign colour black to non-clustered pixels and contrasting colours for clusters
  for (i in 1:nrow(dataframe)){
    
    if(dataframe[i, ]['cluster'] != 0){
      dataframe[i, ]['color'] = colours[as.numeric(dataframe[i, ]['cluster'])]
    } else{
      dataframe[i , ]['color'] = '#000000'
    }
    
  }
  
  return(dataframe)
}
```







## Loading images {.tabset .tabset-pills}

Following article analyzes 2 images using various clustering methods. These are the photographs of bobcat and deer.


### Bobcat
```{r}
#loading image
bobcat = readJPEG("bobcat.jpeg")

#checking dimensions to be sure that it's not a RGB image
dim(bobcat)

#plotting image
bobcat = t(apply(bobcat, 2, rev)) #otherwise the image will be rotated
image(bobcat, col  = grey((0:dim(bobcat)[1])/dim(bobcat)[1]))

```

### Deer
```{r}
#loading image
deer = readJPEG("deer.jpeg")

#checking dimensions to be sure that it's not a RGB image
dim(deer)

#plotting image
deer = t(apply(deer, 2, rev)) #otherwise the image will be rotated
image(deer, col  = grey((0:dim(deer)[1])/dim(deer)[1]))

```

In order to increase the contrast of the image, I am using created function {contrast_image}.
```{r}
#Contrasting image
contrasted_image_bobcat = contrast_image(bobcat, scalar_white = 1.2, scalar_black = 0.1)
contrasted_image_deer = contrast_image(deer, scalar_white = 1, scalar_black = 1.3)
```

**Now it's time to compare clustering methods**

Let's start with CLARA. The intuition behind this algorithm is the same as with PAM, but CLARA works faster with larger datasets, this is why I chose this version to work with images.

## CLARA {.tabset .tabset-pills}

### Bobcat
```{r}
image(bobcat, col  = grey((0:dim(bobcat)[1])/dim(bobcat)[1]))
```

### CLARA with bobcat
```{r}
#CLARA with contrasted image
clara = clara(contrasted_image_bobcat, 2)
colours <- pick_colours(length(unique(clara$cluster)))
img = image(bobcat, col  = colours) # plot in colours
```

### Deer
```{r}
image(deer, col  = grey((0:dim(bobcat)[1])/dim(bobcat)[1]))
```

### CLARA with deer
```{r}
#CLARA with contrasted image
clara = clara(contrasted_image_deer, 3)
colours <- pick_colours(length(unique(clara$cluster)))
img = image(deer, col  = colours) # plot in colours
```

## DBSCAN {.tabset .tabset-pills}

### Bobcat
```{r}
image(bobcat, col  = grey((0:dim(bobcat)[1])/dim(bobcat)[1]))
```

### DBSCAN with bobcat
```{r}
#DB-SCAN with contrasted image
db_scan = dbscan(contrasted_image_bobcat, eps=0.4, MinPts=10) #performing dbscan

df_image = data.frame(contrasted_image_bobcat)
df_image$cluster = db_scan$cluster #adding information about assigned cluster
df_image$color = "" #setting a column for a color

df_image = draw_colours_dbscan(df_image, n_clusters = length(unique(db_scan$cluster))) #sampling colors for clusters
colours = df_image$color 

image(bobcat, col = colours) #drawing image
```

### Deer
```{r}
image(deer, col  = grey((0:dim(bobcat)[1])/dim(bobcat)[1]))
```

### DBSCAN with deer
```{r}
#DB-SCAN with contrasted image
db_scan = dbscan(contrasted_image_deer, eps=0.6, MinPts=3) #performing dbscan

df_image = data.frame(contrasted_image_deer)
df_image$cluster = db_scan$cluster #adding information about assigned cluster
df_image$color = "" #setting a column for a color

df_image = draw_colours_dbscan(df_image, n_clusters = length(unique(db_scan$cluster))) #sampling colors for clusters
colours = df_image$color 

image(deer, col = colours) #drawing image
```

## MEAN SHIFT {.tabset .tabset-pills}

### Bobcat
```{r}
image(bobcat, col  = grey((0:dim(bobcat)[1])/dim(bobcat)[1]))
```

### Mean shift clustering with bobcat
```{r}
#MEAN SHIFT with contrasted image
mean_shift = meanShift(contrasted_image_bobcat, epsilonCluster=0.00052)
df_image = data.frame(contrasted_image_bobcat)
df_image$cluster = mean_shift$assignment #adding information about assigned cluster
df_image$color = ""

df_image = draw_colours(df_image, n_clusters = length(unique(mean_shift$assignment))) #sampling colors for clusters
colours = df_image$color 

image(bobcat, col = colours) #drawing image
```

### Deer
```{r}
image(deer, col  = grey((0:dim(bobcat)[1])/dim(bobcat)[1]))
```

### Mean shift clustering with deer
```{r}
#MEAN SHIFT with contrasted image
mean_shift = meanShift(contrasted_image_deer, epsilonCluster=0.00052)
df_image = data.frame(contrasted_image_deer)
df_image$cluster = mean_shift$assignment #adding information about assigned cluster
df_image$color = ""

df_image = draw_colours(df_image, n_clusters = length(unique(mean_shift$assignment))) #sampling colors for clusters
colours = df_image$color 

image(deer, col = colours) #drawing image
```

## Spectral clustering {.tabset .tabset-pills}

### Bobcat
```{r}
image(bobcat, col  = grey((0:dim(bobcat)[1])/dim(bobcat)[1]))
```

### Spectral clustering with bobcat
```{r}
#SPECTRAL CLUSTERING with contrasted image
spectral = SPECTRAL(contrasted_image_bobcat, k = 4, sigma = 6, graph = FALSE)

df_image = data.frame(contrasted_image_bobcat)
df_image$cluster = spectral$cluster #adding information about assigned cluster
df_image$color = "" #setting a column for a color

df_image = draw_colours(df_image, n_clusters = length(unique(spectral$cluster))) #sampling colors for clusters
colours = df_image$color

image(bobcat, col = colours)
```

### Deer
```{r}
image(deer, col  = grey((0:dim(bobcat)[1])/dim(bobcat)[1]))
```

### Spectral clustering with deer
```{r}
#SPECTRAL CLUSTERING with contrasted image
spectral = SPECTRAL(contrasted_image_deer, k = 2, sigma = 2.5, graph = FALSE)

df_image = data.frame(contrasted_image_deer)
df_image$cluster = spectral$cluster #adding information about assigned cluster
df_image$color = "" #setting a column for a color

df_image = draw_colours(df_image, n_clusters = length(unique(spectral$cluster))) #sampling colors for clusters
colours = df_image$color

image(deer, col = colours)
```